---
title: "lab_02"
author: "derek willis"
date: "8/09/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings
```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
library(lubridate)
```

## Load Data

Let's keep working with the slice of WinRed contribution data from West Virginia.  

```{r}
wv_winred_contribs <- read_rds("data/lab_02.rds")

head (wv_winred_contribs)
```

## Examine Data

## Answer questions

**Q1** Create a new column that divides the contribution amounts into the following options - less than $10, $10-$25, $26-$50, $51_$100, $101-$500, $501 and up - and calculates how many contributions are in each group. How would you describe the distribution of contributions in a sentence?

A1. 5672 contributions were given between $10 and $25. This was the most. The second most contributions given were less than $10 with 4638 contributions.Only 38 people donated more than $501.  

```{r}
wv_winred_categories <- wv_winred_contribs %>%
  mutate(
    amount_category = case_when(
        amount < 10 ~ "Less than $10",
        amount >= 10 & amount <= 25 ~ "Between $10 and $25",
        amount > 25 & amount <= 50 ~ "Between $26 and $50",
        amount > 50 & amount <= 100 ~ "Between $50 and 100",
        amount > 100 & amount <= 500 ~ "Between $101 and $500",
        amount > 501  ~ "More than $501"
      )
  ) %>% 
  group_by(amount_category) %>% 
  summarize(count_contribs = n()) %>% 
  arrange(desc(count_contribs))

```


```{r}
```
Q2. Which West Virginia donors made the most contributions of less than $10? You'll need to use a combination of columns to get unique donors.  
A2. Rev Dr. John made the most contributions less than $10. 

wv_winred_categories <- 

```{r}
wv_winred_contribs %>%
  mutate(
    amount_category = case_when(
        amount < 10 ~ "Less than $10",
        amount >= 10 & amount <= 25 ~ "Between $10 and $25",
        amount > 25 & amount <= 50 ~ "Between $26 and $50",
        amount > 50 & amount <= 100 ~ "Between $50 and 100",
        amount > 100 & amount <= 500 ~ "Between $101 and $500",
        amount > 501  ~ "More than $501"
      )
  ) %>% 
  filter (amount <10) %>% 
group_by(last_name, first_name, address_one) %>% 
  summarize(count_contribs = n()) %>% 
  arrange(desc(count_contribs))

```


Q3. What questions do the results of Q2 raise in your mind? How could you use the data to explore them? No code here, just your questions.
A3. I'm curious as to which candidate that Rev Dr. John donated to and what makes him want to support that particular platform. 293 are a lot of contributions.Based on his title, I assume that he works for a church. We could look further into the church and his connection to the particular campaign that he's donating to. 

Q4. Which month saw the most contributions?
A4. June saw the most contributions with 4636 contributions. 

```{r}
wv_winred_contribs %>%
  mutate(Month = floor_date(date, "month")) %>% 
group_by(Month) %>% 
  summarize(count_contribs = n()) %>% 
  arrange(desc(count_contribs))

```

Q5. How many contributions of $2,900 - the individual maximum donation - were there, and who did they go to? How many of the recipients are from West Virginia? Do some googling to find out.
A5. 
There were six contributions of 2,900. The contributions went to Carol for Congress, McKinley for Congress, Palazzo for Congress, and Zinke for Congress.

How many of the recipients are from West Virginia?:2, both Carol for Congress and McKinley for Congress are from West Virginia.

```{r}
wv_winred_contribs %>%
 filter (amount == 2900) %>% 
  arrange(date)

```
