---
title: "Class Reference"
author: "Student name"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Consider this a personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.


## Things I Learned on Jan. 27

The command to set a working directory is setwd(). For example:

```{r}
setwd("~/Documents/GitHub/data_journalism_2022_spring")
```

```{r}
install.packages("tidyverse")
library(tidyverse)
```


### Summarizing

I need to use group_by and summariz. Here's an example of grouping by county and calculating counts, sum and other descriptive statistics.


```{r}
ppp_maryland_loans %>%
  group_by(project_county_name) %>%
  summarise(
    count_loans = n(),
    total_loans_amount = sum(amount),
    mean_loan_amount = mean(amount),
    median_loan_amount = median(amount),
    min_loan_amount = min(amount),
    max_loan_amount = max(amount)
  ) %>%
  arrange(desc(max_loan_amount))
```


Pre-lab 02 Notes:

```{r}
head(primary_18)
summary(primary_18)
colnames(primary_18)
glimpse(primary_18)
```

Head- Puts data in a chart
Summary- Provides an overall summart of the data from each cataegory from the chart (Max, Min, Median, mean, ect...)
Colnames- shows each of the column names from the chart.
Glimpse- Shows data type along with the values from the different columns in the chart. 

Mutate:adds new column base on exisiting columns. Populate data in new column using other columns. 
 Ex:
 
```{r}
primary_18_%>%
  select(office, district, name_raw, party, jurisdiction, election_day, votes) %>% 
  mutate(
  percent_election_day = election_day/votes
)
```
 Select: Means I only want to see a certain number of rows. 
 
 Ordering the columns using arrange:
 
```{r}
primary_18 %>%
  select(office, district, name_raw, party, jurisdiction, election_day, votes) %>% 
  mutate(
  percent_election_day = (election_day/votes)*100
)  %>% arrange(desc(percent_election_day))
```
 
You'll notice that there's a mix of styles: "Baltimore" and "BALTIMORE" for example. R will think those are two different cities, and that will mean that any aggregates we create based on city won't be accurate.

So how can we fix that? Mutate - it's not just for math! And a function called `str_to_upper` that will convert a character column into all uppercase. Now we can say exactly how many donations came from Baltimore (I mean, of course, BALTIMORE).

```{r}
standardized_maryland_cities <- maryland_cities %>%
  mutate(
    upper_city = str_to_upper(city)
)
```

```{r}
standardized_maryland_cities %>% 
  group_by(upper_city) %>% 
  summarize(total = sum(sum), total_count = sum(count))
```

Str_to_lower- Converts a string to lowercase (Letters of each word in a string are lowercased)
Str_to_title- Converts the first letter of each word in a string to uppercase and the rest to lower case. 

Mutate is even more useful when combined with some additional functions. Let's focus on individual contributions from Maryland donors via WinRed; we'd like to group their donations by amount into one of four categories:

1. Under $100
2. $101-$499
3. $500-$2,900
4. More than $2,901

Mutate can make that happen by creating a new column and putting in a category value _based on the amount_ of each record. First, let's load some individual contributions:

Case When Example:

```{r}
maryland_winred_categories <- maryland_winred %>%
  mutate(
    amount_category = case_when(
        amount < 100 ~ "Less than $100",
        amount >= 100 & amount <= 500 ~ "Between $100 and $500",
        amount > 500 & amount <= 2900 ~ "Between $501 and $2900",
        amount > 2900 ~ "More than $2,900"
      )
  )
head(maryland_winred_categories)
```

```{r}
maryland_winred_categories %>% 
  group_by(amount_category) %>% 
  summarize(total_amount = sum(amount)) %>% 
  arrange(desc(total_amount))
```


Mutate is there to make your data more useful and to make it easier for you to ask more and better questions of it.


****Comparission operators:

We need to filter all records equal to "Prince George's". The comparison operators in R, like most programming languages, are == for equal to, != for not equal to, > for greater than, >= for greater than or equal to and so on.


From pre lab 01 variable example: 
Name of varaible, left arrow dash symbol, and then the value. 
number <- -2  

What does nrow mean? 


```{r}
or_prince_georges <- primary_18 %>% filter(jurisdiction == "Prince George's" | office == "Governor / Lt. Governor")

nrow(or_prince_georges)
```

So there's 355 rows that are EITHER in Prince George's OR are governor results. OR is additive; AND is restrictive.

Run the following code and describe the change in the expenditure_date column.

```{r}
maryland_expenses <- maryland_expenses %>% mutate(expenditure_date=mdy(expenditure_date))

head(maryland_expenses)
```


Lubridate has functions for basically any type of character date format: mdy, ymd, even datetimes like ymd_hms.

Fortunately, `readr` anticipates some date formatting and can automatically handle many of these issues (indeed it uses lubridate under the hood). When you are importing a CSV file, be sure to use `read_csv`, not `read.csv`.

And Then symbol----- Shift Command M (One of these Two) (Example Below)
%>% 

Review:
**Filter: you use it when you just want to see data that meets a certain criteria.
**Select-Just to see certain columns. (Can also say that you want to get everything except those by putting minus sign infront of those names i.e.-committee_name)

rm (data frame name from the environment)- How you remove something from environment. 

Floordate, takes date and puts it near the nearest integer value. 

Cleaning Names Funtion: This function allows you to fix the names of columns.


cleaned_conowingo <- conowingo %>%
  clean_names() %>% 
  rename(linenumber = x1_linenumber) %>% 
  mutate(amount = as.numeric(amount)) %>% 
  distinct() %>%
  mutate(zip = str_sub(zip, start=1L, end=5L))
  
  distinct () - lob 03,lets you keep just distinct value of a function. Used when dealing with duplicates. 
  
county_voters_combined <- bind_rows(list(county_voters_2016, county_voters_2018, county_voters_2020))
View(county_voters_combined)

^^^Combining similar columns into one chart for different data frames. 
As long as you have the same names, order does not matter. 

maryland_population_with_voters <- maryland_population %>% left_join(county_voters_2020)

^^^Left Join example. Adding smaller data coulmn to bigger set. There needs to be a common colmun in both.

maryland_population_with_voters_and_demographics <- maryland_population_with_voters %>% left_join(maryland_demographics, by=c("COUNTY"="NAME"))  

county_voters_combined <- bind_rows(list(county_voters_2016, county_voters_2018, county_voters_2020))
View(county_voters_combined)

list- There's a list of data frames comply.

To add everything in a column- sum(namedofdataframe$nameofcolumn)


# cleaning function
cleaned_conowingo <- conowingo %>%
  clean_names()

# display the cleaned dataset
cleaned_conowingo


# cleaning function
cleaned_conowingo <- conowingo %>%
  clean_names() %>% 
  rename(linenumber = x1_linenumber) %>% 
  mutate(amount = as.numeric(amount)) %>% 
  distinct()
  

# display the cleaned dataset
cleaned_conowingo

%>% %>% 


