---
title: "State Profile Final"
output: html_document
date: "2022-11-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r cars}

```

```{r}
library('tidyverse')
library('janitor')
library('lubridate')
```



My state: MI (Michigan)

Democratic Party ID: C00031054

Republican Party ID: C00041160

Data: MIindividualContri.csv 

**Using receipts because he said that the individual contributions would be in the receipts. Check back on Dec. 8 for updated data. 
For spending, I used the disbursements data. 

```{r}
MIDemRecipts <- read_csv("StatesData/MiDemAllRecipts.csv")
MIRepRecipts <- read_csv("StatesData/MIRepAllRecipts.csv")
MiRepSepnding <- read_csv("StatesData/MIRepSpending.csv")
MiDemSpending <- read_csv("StatesData/MiDemSpending.csv")
```
-How can you describe (broadly) the party’s donors? Where do the parties get their money and from what kinds of donors? Do they raise a lot in small amounts or from out of state donors?

-What did the party spend its money on? You may need OpenRefine to help you standardize the purpose of spending. Can you tell much about the strategy behind the spending from the data?

-When you compare the two parties, where are the differences and similarities?

-What about either the donors or the spending prompts further questions from you?

Recipts:
committee_name_2
report_year
entity_type_desc- Filter for individual donnations. 
contributor_first_name
contributor_last_name
contributor_street_1
contributor_city
contributor_state
contributor_employer
contributor_occupation
contribution_receipt_date
contribution_receipt_amount

#Organizing and cleaning Michigan Democratic recipts:

```{r}
CleanMIDemR <- MIDemRecipts %>% 
  clean_names() %>% 
  select(committee_name_2,contributor_name,report_year,entity_type_desc,contributor_street_1,contributor_city,contributor_state,contributor_employer,contributor_occupation, contribution_receipt_date,contribution_receipt_amount ) %>% 
  group_by(contribution_receipt_amount) %>% 
  arrange(desc(contribution_receipt_amount))
```

#Who made the most contributions to the Michigan Democratic Party:

```{r}
MostContribsDem <- CleanMIDemR %>% 
  group_by(contributor_name) %>% 
summarise(count=n()
) %>%
  arrange(desc(count))
```

#Who donated the most money to the Michgan Democratic Party?

```{r}
MoneyMostDem <- CleanMIDemR %>% 
  group_by(contributor_name) %>% 
  summarise(total= sum(contribution_receipt_amount)
) %>%
  arrange(desc(total)) %>% 
  distinct()
```

#Which individual donated the most money to the Democratic Party?

```{r}
DemTopIndiContrib <- CleanMIDemR %>% 
  filter(entity_type_desc == "INDIVIDUAL") %>% 
  group_by(contributor_name) %>% 
  summarise(total= sum(contribution_receipt_amount)
) %>%
  arrange(desc(total)) %>% 
  distinct()
```


#Organizing and Cleaning Michigan Republican recipts:

```{r}
CleanMIRepR <- MIRepRecipts %>% 
  clean_names() %>% 
  select(committee_name_2,contributor_name,report_year,entity_type_desc,contributor_street_1,contributor_city,contributor_state,contributor_employer,contributor_occupation, contribution_receipt_date,contribution_receipt_amount ) %>% 
  group_by(contribution_receipt_amount) %>% 
  arrange(desc(contribution_receipt_amount))
```

#Who made the most contributions Michigan Republican Party?

```{r}



```

#Who donated the most money to the Michigan Republican Party?

```{r}

```


```{r}

```


#Just Democratic Individual Contributions from recipts

```{r}
DemIndiContribs <- CleanMIDemR %>% 
  filter(entity_type_desc == "INDIVIDUAL")
```

#Just Republican Individual Contributions from recipts

```{r}
RepIndiContribs <- CleanMIRepR %>% 
    filter(entity_type_desc == "INDIVIDUAL")
```









#Cleaning and organizing spending Democratic Data

I don't know what fec election columns mean.

```{r}
CleanMiDemSpend <- MiDemSpending %>% 
  clean_names() %>% 
  select(committee_name,report_year,entity_type_desc,recipient_name,recipient_state,beneficiary_committee_name, disbursement_type_description, disbursement_description, disbursement_date, disbursement_amount, fec_election_type_desc,fec_election_year) %>% 
  group_by(disbursement_amount) %>% 
  arrange(desc(disbursement_amount))
```

#Cleaning and organizing republican spending data.

```{r}
CleanMiRepSpend <- MiRepSepnding %>% 
  clean_names() %>% 
  select(committee_name,report_year,entity_type_desc,recipient_name,recipient_state,beneficiary_committee_name, disbursement_type_description, disbursement_description, disbursement_date, disbursement_amount, fec_election_type_desc,fec_election_year) %>% 
  group_by(disbursement_amount) %>% 
  arrange(desc(disbursement_amount))
```



