---
title: "State Profile Final"
output: html_document
date: "2022-11-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library('tidyverse')
library('janitor')
library('lubridate')
library('refinr')
```


My state: MI (Michigan)

Democratic Party ID: C00031054

Republican Party ID: C00041160

Data: MIindividualContri.csv 

**Using receipts because he said that the individual contributions would be in the receipts. Check back on Dec. 8 for updated data. 
For spending, I used the disbursements data. 

```{r}
MIDemRecipts <- read_csv("StatesData/MiDemAllRecipts.csv")
MIRepRecipts <- read_csv("StatesData/MIRepAllRecipts.csv")
MiRepSepnding <- read_csv("StatesData/MIRepRefSpending-csv.csv")
MiDemSpending <- read_csv("StatesData/MiDemRefSpending-csv.csv")
```

***How can you describe (broadly) the party’s donors? Where do the parties get their money and from what kinds of donors? Do they raise a lot in small amounts or from out of state donors?

Answer: The Michigan Democratic Party received the most money in donations with a total of $10382351.The Michigan Republican Party received $2520106. That's a difference of $7,862,245.

The Michigan Democratic Party received most from Political Party Committees (38%) and individual donors (27%). The Michigan Republican Party received most of its money from Individuals (99.9%) and Political Action Committees (0.06%). It surprised me that the Republican Party did not receive more money from political action committees and that it received none from any other organizations. 

Half of the top 20 organizations that gave the most money to the Democratic Party are from the campaigns of other Michigan government officials. Democratic Parties in Ohio, Utah, and Indiana also significantly contributed. 

Most of the Democratic Party's money came from Washington DC ($4717282.54) with the Democratic National Committee Contributing most of that. Michigan, Massachusetts, California and Ohio round out the top five. The Michigan Republican Party receives most of its money from the state of Michigan at $1679098.75.Florida gave the party the second most in terms of money at $273,476.29. California, Colorado, and Texas round out the top five.

The Democratic Party received most of its contributions in $30 dollar donations. 1658 contributions came in the amount of $30. 1311 Democratic contributions came in the amount of $25. 1115 contributions came in the amount of $100. Most of the Republican Party's donations came in the amount of $25. 1322 contributions were made in that amount. 1085 contributions were made in the amount of $10 and 656 were made in the amount of $5.The Republican Party's most popular donation amounts were less than the Democrats. 

In terms of the most individual donors, the top individual donor for the Democratic Party was Merle Catherine Chambers from Colorado, who donated $27200 in three contributions. Chambers is a lawyer, investor, and philanthropist that has a history of donating to other Democratic campaigns in Colorado, New Hampshire, and Missouri. She gave money to organizations connected to Bill Clinton in 1992. I think a profile story on Catherine Chamber could be interesting. She could talk about her rationale for supporting the party, and discuss the state of it.

Betsy Devos, former United States Secretary of Education from 2017 to 2021, donated $20,000 to the Republican Party. She is from Michigan. Her husband, Richard Devos JR, and five other members of their family each donated $20,000. They each gave two contributions of $10,000. David Devos only donated $400.

Just looking at the Republican Contributions data, there appears that more groups of families contributed to the Republican Party than in the Democratic contribution dataset.

***What did the party spend its money on? You may need OpenRefine to help you standardize the purpose of spending. Can you tell much about the strategy behind the spending from the data?

A: Ultimately, the Republican Party spent less money than the Democratic Party. Michigan's Republican Party spent $1427744 compared to the Democratic Party that spent $5026673. That's a $3,598,929 difference. 

The top overall transaction by the Democratic Party was $816204.67 given to The Strategy Group for Absentee Ballot Mailer. The Democratic Party also spent the most money at The Strategy Group, spending $866204.67 total. $50000 of that total was spent on direct mail voter protection. It's clear that the Democratic Party in Michigan has continued to focus on mail-in voting. During the pandemic, there was a huge emphasis on mail-in voting because many voters did not feel comfortable going to the voting booths. Ultimately, mail-in voting provides more people with the chance to vote, which is why the Democratic Party is investing so much into it. 

The Michigan Republican Party gave most of its money to itself, spending $287673.18. This could just be that party transfering money from its federal account to it state account or vice versa.I found it interesting that the top choice for advertising for the Michigan Republican Party was online. They party spent $149,195.29 on online advertising with the company IMGE. It's also interesting that the state's Republican Party spent more than $40,000 on Voter ID registration software from Campaign Sidekick. Does this have to do with the national Republican Party's lack of trust in the voting system? 

The Democratic Party spent a lot of money on GOTV (Get out the vote) for advertising. The recipient of this transaction was Facebook. I believe that Facebook advertising was the top choice for Democrats. So, both parties invested heavily in online advertising. 

Both parties invested heavily in mail, as it is a top 5 expense for both parties. 

***What about either the donors or the spending prompts further questions from you?

I am curious as to why other states Democratic Parties decided to specifically donate to Michigan. The Ohio Democratic Party is in the top 5 of money donated to the party with $435000. Indiana and Utah's Democratic Parties are in the top 20 of money donated to Michigan's Democratic Party.

Is there a reason why more families tend to donate more money to the Republican Party than in the Democratic Party?

Why aren't more political action committees supporting the Republican Party? Why are there no other organizations contributing to the party?

Is the Michigan Republican Party's $42,000 investment in Voter ID registration software a reflection of the national Republican Party's lack of trust in the voting system? 

Is there a reason why the most popular donations to Michigan's Republican Party tend to be lower?

How does Catherine Chambers decide which Democratic Parties she wants to donate to? Is there some type of pattern to her donations. 

#Organizing and cleaning Michigan Democratic recipts:

```{r}
CleanMIDemR <- MIDemRecipts %>% 
  clean_names() %>% 
  select(committee_name_2,contributor_name,report_year,entity_type_desc,contributor_street_1,contributor_city,contributor_state,contributor_employer,contributor_occupation, contribution_receipt_date,contribution_receipt_amount ) %>% 
  group_by(contribution_receipt_amount) %>% 
  arrange(desc(contribution_receipt_amount))
```

#What entity made the most contributions to the Michigan Democratic Party:

From January 01, 2021 through October 19, 2022, ActBlue Technical Services leads all contributors with 4744 donations. According to the organization's website,  Democratic campaigns use Actblue to "to create online contribution forms, securely process donations, and track their contributions from grassroots supporters." From my understanding, it's not the organization itself that's donating the money, but its money from individual contributors or organzations that's processed using ActBlue services. The organization does not support any particular candidates. Most people who donate using this service give about $30 to $25.However, there were 20 donations using the service of $10,000. I guess that contributors who use this service can remain anonymous as the names of these contributors is not mentioned in this data. 

Other than the Democratic National Committee, the other entities who made the most contributions are individuals: Ellen Kerr, Jeffery Smith, and Joseph Hess. Smith gave 56 contributions for $3,750, while Hess and Kerr donated a little over $1,000. Hess is a attorney and owner at his own healthcare lawfirm in Grand Rapids Michigan. His donations make sense given the Democratic's stance on abortion rights as well as Michigan Democratic Gov. Whitmer.When I search for Jeffery Smith, I see that there's a partner at a law firm that provides healthcare services in Michigan and a doctor of medicine at the University of Michigan Health with the same name. I think that "Ellen Kerr" could be Lee Ellen Kerr, is the Vice President of an nonprofit called Antrim County High Tea for Breast Cancer Prevention. 

```{r}
MostContribsDem <- CleanMIDemR %>% 
  group_by(contributor_name,entity_type_desc) %>% 
summarise(count=n()
) %>%
  arrange(desc(count))
```

#Filtering specifically for ActBlue:

```{r}
ActBlue <- CleanMIDemR %>% 
  clean_names() %>% 
  select(committee_name_2,contributor_name,contribution_receipt_date,contribution_receipt_amount ) %>% 
  group_by(contribution_receipt_amount) %>% 
  filter(contributor_name == "ACTBLUE TECHNICAL SERVICES") %>% 
  arrange(desc(contribution_receipt_amount))
```

```{r}
ActBlueContribAmounts <- ActBlue %>% 
   group_by(contribution_receipt_amount) %>% 
summarise(count=n()) %>%
  arrange(desc(contribution_receipt_amount))
```

#Filtering Ellen, Kerr

```{r}
EllenK <- CleanMIDemR %>% 
  clean_names() %>% 
  select(committee_name_2,contributor_name,contribution_receipt_date,contribution_receipt_amount ) %>% 
  group_by(contribution_receipt_amount) %>% 
  filter(contributor_name == "KERR, ELLEN") %>% 
  arrange(desc(contribution_receipt_amount))

EllenKContribAmounts <- EllenK %>% 
   group_by(contribution_receipt_amount) %>% 
summarise(count=n()) %>%
  arrange(desc(contribution_receipt_amount))
```

#Who donated the most money to the Michgan Democratic Party?
Overall, most of the money donated to the Michigan Democratic Party came from the Democratic National Committee at $2588969.12 and ActBlue at $928739.76. Michigan U.S. Senator Debbie Stabenow's fund donated the third most money to the Democratic Party at $554587.83. About half of the top 20 organizations that gave the most money are from the campaigns of other Michigan government officials. I am curious as to why other states Democratic Parties decided to specifically donate to Michigan. The Ohio Democratic Party that is in the top 5 with $435000. Indiana and Utah's Democratic Parties are in the top 20 of donations to Michigan's Democratic Party. 

```{r}
MoneyMostDem <- CleanMIDemR %>% 
  group_by(contributor_name,contributor_state) %>% 
  mutate(contributor_name = case_when(contributor_name == "DNC" ~ "DEMOCRATIC NATIONAL COMMITTEE",contributor_name == "DNC SERVICES CORP/DNC" ~ "DNC SERVICES CORP.",TRUE ~ contributor_name)) %>% 
  summarise(total= sum(contribution_receipt_amount)
) %>%
  arrange(desc(total)) %>% 
  distinct()
```

#Which individual donated the most money to the Democratic Party?

In terms of individual donors, Merle Catherine Chambers from Colorado donated the most money to Michigan's Democratic Party. She donated $27200 in three contributions. Two $10,000 donations and one $7,200 donation. According to the website influencewatch.org, Chambers is lawyer, investor, and former oil and natural gas CEO. From 1980 to 1997, she was the founding President and CEO of Axem Resources. She is also a philanthropist who founded the Merle Chambers Fund, which donates to abortion, social justice, education, economic security for women, and other advocacy causes supported by the Democratic Party. She has a history of donating to other Democratic campaigns of candidates in Colorado, New Hampshire, and Missouri. She also donated money to organizations connected to Bill Clinton in 1992. 

The third leading donor to the Michigan Democrats was Robert Price, CEO of PriceSmart, which is like a warehouse store company. According to a 2022 article on Wallmine, his net worth is about $234 million. His wife, Allison gave the second-most money to Michigan's Democratic Party. They're both from California.

The only two people in the top five in money donated to the state's Democratic Party that are from Michigan are James Kelmer, who is the VP of Intellectual Property Bussiness Strategy at Strkyer Corp, and James Sanford Offield.

```{r}
DemTopIndiContrib <- CleanMIDemR %>% 
  filter(entity_type_desc == "INDIVIDUAL") %>% 
  group_by(contributor_name, contributor_state) %>% 
  summarise(total= sum(contribution_receipt_amount)
) %>%
  arrange(desc(total)) %>% 
  distinct()
```

```{r}
Merle <- CleanMIDemR %>% 
  filter(contributor_name == "CHAMBERS, MERLE CATHERINE")
JamesKel <-CleanMIDemR %>% 
  filter(contributor_name == "KEMLER, JAMES")
JamesSan <- CleanMIDemR%>% 
  filter(contributor_name == "OFFIELD, JAMES SANFORD")
```

```{r}
Robert <- CleanMIDemR %>% 
  filter(contributor_name == "PRICE, ROBERT")
```

#Which type of entity is donating most of the money for the Democratic Party?

The Michigan Democratic Party received most of their money from Political Party Committees. More than 37% of the total amount of money donated to the party came from Political Party Committees. Individual donations made up about 26% of the total contributions. Political Action Committees made up about 17% of the contributions. 

```{r}
DonatingMostDem <- CleanMIDemR %>% 
  group_by(entity_type_desc) %>% 
  summarise(total= sum(contribution_receipt_amount)
) %>%
   mutate(percent = (total/10382351)*100) %>% 
  arrange(desc(total))
```

#Which type of entity made the most number donations to the Democratic Party? (Not amount of money) -Why are there more categories for the Republicans than the Democrats?

Individual donors made the most contributions at 5389. Organizations made second most contributions at 4824. However, 4744 of those contributions were from ActBlue. Some of the other organizations that were accounted in this contribution count were National Capital Bank (13 contributions), The Sbam Plan (10 contributions), Paychex (9 contributions). Ironically, Political Action Committees came in second to last in contributions at 78. This makes sense considering that their contributions tend have more money attached to them. Campaign Committees made the fewest number of contributions with 24.

```{r}
EntityDonationDem <- CleanMIDemR %>% 
  group_by(entity_type_desc) %>% 
summarise(count=n()
) %>%
  arrange(desc(count))
```

#What was the was the smallest amount given? --- Does negative money mean organization had to take their money back because they contributed over the limit?

```{r}
BottomAmountDem <- CleanMIDemR %>% 
  group_by(contribution_receipt_amount) %>% 
  arrange((contribution_receipt_amount))
```

***Could filter contribution amounts for a certain price number to see if most of the donations are small or bigger. 

#Organizing and Cleaning Michigan Republican recipts:

```{r}
CleanMIRepR <- MIRepRecipts %>% 
  clean_names() %>% 
  select(committee_name_2,contributor_name,report_year,entity_type_desc,contributor_street_1,contributor_city,contributor_state,contributor_employer,contributor_occupation, contribution_receipt_date,contribution_receipt_amount ) %>% 
  group_by(contribution_receipt_amount) %>% 
  arrange(desc(contribution_receipt_amount))
CleanMIRepR
```

#What entity type is donating the most money to the Republican Party? 

Individuals donated the most money to Michigan's Republican Party at $2517661.37, which is more than 99% of the total donations that the party received. Political Action Committees donated $1556.92 and Other committees donated $887.51. It surprised me that individuals made up such a large percentage. It's also surprising that organizations and other campaigns did not make any contributions.

```{r}
DonatingMostRep <- CleanMIRepR %>% 
  group_by(entity_type_desc) %>% 
  summarise(total= sum(contribution_receipt_amount)
) %>%mutate(percent = (total/2520106)*100) %>% 
  arrange(desc(total))
```

#What entity gave the most contributions Michigan Republican Party?

Individuals made up nine of the top ten with the highest number of contributions given to the party. This differs from the Democratic Party, where two of the top five in contributions are a political party committee (Democratic National Committee) and a organization (ActBlue).

Karl Szymanski from Michigan gave the Michigan Republican Party the most contributions at 76. Two of his donations were $100, one was $75, and the rest were $50 and under. The data says that he is retired. Karl and Ronald Davis are the only two people to donate more than 30 contributions. The others in the data set are below 30. 

Winred, the Republican's version of ActBlue, comes in a number 10 on most contributions with 23. I'm curious as to why this platform is not as popular as ActBlue. Are committees not promoting it as much?

```{r}
MostContribsRep <- CleanMIRepR %>% 
  group_by(contributor_name) %>% 
summarise(count=n()
) %>%
  arrange(desc(count))

KarlContribs <- CleanMIRepR %>% 
  clean_names() %>% 
  select(committee_name_2,contributor_name,contribution_receipt_date,contribution_receipt_amount ) %>% 
  group_by(contribution_receipt_amount) %>% 
  filter(contributor_name == "SZYMANSKI, KARL MR.") %>% 
  arrange(desc(contribution_receipt_amount))
```

#Which type of entity made the most number donations to the Republican Party? (Not amount of money)

Individuals gave the most contributions with 5966. Political Action Committee gave 15 contributions. Other committees provided 8 contributions. Not as many contributions came from political action committees as they did from the Democratic Party. The Republican Party only received 15 contributions from political action committees, all of which were from Winred. 

```{r}
EntityDonationRep <- CleanMIRepR %>% 
  group_by(entity_type_desc) %>% 
summarise(count=n()
) %>%
  arrange(desc(count))
```

#Who donated the most money to the Michigan Republican Party?

Nine people in the top 10 tied in donating the most money. Each of them gave $40,000. They're each individual donors. 

Timothy Easterwood from Michigan is a CEO at a company called BizFleets, which is a fleet management service that reduces the cost of owning and operating work vehicles for businesses.

Jay Fishman is the President and Founder of Jay A Fishman Investment Counsel. 

Herbert Henkle, who is on the Board of Directors at Ingersoll-Rand Company, and his wife Linda Henkle. 

Betsy Devos, former United States Secretary of Education from 2017 to 2021, donated $20,000. She is from Michigan. Her husband, Richard Devos JR, and five other members of their family each donated $20,000. They each gave two contributions of $10,000. David Devos only donated $400.

Just looking at the Republican Contributions data, there appears that more groups of families contributed than in the Democratic contribution dataset. 

```{r}
MoneyMostRep <-CleanMIRepR  %>% 
  group_by(contributor_name,contributor_state) %>% 
  summarise(total= sum(contribution_receipt_amount)
) %>%
  arrange(desc(total)) %>% 
   distinct()
```

#Just Democratic Individual Contributions from recipts

```{r}
DemIndiContribs <- CleanMIDemR %>% 
  filter(entity_type_desc == "INDIVIDUAL")
```

#Just Republican Individual Contributions from recipts

```{r}
RepIndiContribs <- CleanMIRepR %>% 
    filter(entity_type_desc == "INDIVIDUAL")
```

#What state donated the most to each party?

The Michigan Democratic Party received most of its money from the Washington, D.C. at $4717282.54. Michigan gve the state's party the second most at $2415077.67. Massachusetts, California, and Ohio all round out the top five.

The Michigan Republican Party receives most of its money from the state of Michigan at $1679098.75.Florida gave the party the second most in terms of money at $273,476.29. California, Colorado, and Texas round out the top five. 

```{r}
StateDem <- CleanMIDemR %>% 
  group_by(contributor_state)%>% 
  summarise(total= sum(contribution_receipt_amount)
) %>%
  arrange(desc(total))
```

```{r}
StateRep <-CleanMIRepR %>% 
  group_by(contributor_state)%>% 
  summarise(total= sum(contribution_receipt_amount)
) %>%
  arrange(desc(total))
```

#What is the total amount of money donated to each party?

The Michigan Democratic Party received a total of $10382351.The Michigan Republican Party received $2520106. That's a difference of $7,862,245.

```{r}
TotalContirbDem <- CleanMIDemR %>% 
  group_by(committee_name_2) %>% 
  summarise(total_amount = sum(contribution_receipt_amount))

TotalContribRep <- CleanMIRepR %>% 
  group_by(committee_name_2) %>% 
  summarise(total_amount = sum(contribution_receipt_amount))
```

#Are most donations in small chunks or big per party? 

Most of the donations for the Michigan Democratic Party came in smaller chunks, with 1658 donations of $30. There were 1311 donations of $25. There were 1115 donations of $100 and 915 of $10. The highest contribution amount received for the Democratic Party was $37,0000.

A similar trend is true for the Michigan Republican Party. The party received 1322 donations of $25. 1085 donations were recieved of $10. This is followed by 656 $5 donations. So, most of the contributions for the Republicans are coming in at lower dollar amounts than for the state's Democratic Party. The Democratic party's most popular contribution amount is higher. The highest contribution amount received for the Republican Party was $9950. 

```{r}
DemChunks <- CleanMIDemR %>% 
  group_by(contribution_receipt_amount) %>% 
summarise(count=n()) %>%
  arrange(desc(count))

RepChunks <- CleanMIRepR %>% 
  group_by(contribution_receipt_amount) %>% 
summarise(count=n()) %>%
  arrange(desc(count))
```

#SPENDING 

#Cleaning and organizing spending Democratic Data

```{r}
CleanMiDemSpend <- MiDemSpending %>% 
  clean_names() %>% 
  select(committee_name,report_year,entity_type_desc,recipient_name,recipient_state,beneficiary_committee_name, disbursement_type_description, disbursement_description, disbursement_date, disbursement_amount, fec_election_type_desc,fec_election_year, or_disbursement_description) %>% 
  group_by(disbursement_amount) %>% 
  mutate(disbursement_description = case_when(disbursement_description == "GOTV EVENT ENTERTAINMENT" ~ "GOTV", disbursement_description == "PAYROLL PROCESSING FEES" ~ "PAYROLL FEES", disbursement_description == "GOTV RALLY" ~ "GOTV", disbursement_description == "STAC SERVICES" ~ "STAC LAB SERVICES", disbursement_description == "GOTV ADS - NOT CANDIDATE RELATED" ~ "GOTV", TRUE ~ disbursement_description)) %>% 
  arrange(desc(disbursement_amount))
```

#Who did most of the Democratic money go to? What was the reason of the transaction?

The top overall transaction by the Democratic Party was $816204.67 given to The Strategy Group for Absentee Ballot Mailer. The Democratic Party also spent the most money at The Strategy Group, spending $866204.67 total. $50000.0 of that total was spent on direct mail voter protection. The Strategy Group has previously served as the direct mail strategist for President Obama during his two Presidential campaigns. The company has also been involved in the election of several Democratic members of congress and several Democratic Gubernatorial candidates. The party also spent more money in terms on mail at Mission Control Inc. and Gumbiner and Davies Communications. These companies were also in the top that received the most money from the state's Democratic Party. The party also spent a lot of money, $188958.00 total, on voter registration with SHAWMUT SERVICES LLC. 

```{r}
TopRecieverDem <- CleanMiDemSpend %>% 
  group_by(recipient_name,disbursement_description, or_disbursement_description) %>% 
  summarise(total= sum(disbursement_amount)
) %>%
  arrange(desc(total)) %>% 
  distinct()

TopDemNoDesc <- CleanMiDemSpend %>% 
  group_by(recipient_name) %>% 
   summarise(total= sum(disbursement_amount)
) %>%
  arrange(desc(total)) %>% 
  distinct()

```

#What transaction type was most of the money spent on for the Democratic Party?--

About 27% of the Michigan Democratic Party's money was just spent on the salary of it's employees. About 16% and 15% respectively were spent on absentee ballots and other mail. About 10% was spent on payroll taxes.

```{r}
TransTypeDem <- CleanMiDemSpend %>% 
  group_by(disbursement_description) %>% 
  summarise(total= sum(disbursement_amount)
) %>% mutate(percent = (total/5026673)*100) %>% 
  arrange(desc(total)) %>% 
  distinct()
```


#Cleaning and organizing republican spending data.

```{r}
CleanMiRepSpend <- MiRepSepnding %>% 
  clean_names() %>% 
  select(committee_name,report_year,entity_type_desc,recipient_name,recipient_state,beneficiary_committee_name, disbursement_type_description, disbursement_description, disbursement_date, disbursement_amount, fec_election_type_desc,fec_election_year, or_disbursement_description) %>% 
  group_by(disbursement_amount) %>% 
  mutate(disbursement_description = case_when(disbursement_description == "CC Processing Fee" ~ "CC Processing Fees",disbursement_description == "FOOD FOR FUNDRAISING MEETING- NON CAND." ~ "FOOD", disbursement_description == "FOOD FOR MEETING- VOTER ID/NON CAND." ~ "FOOD",disbursement_description == "FOOD FOR VOLUNTEERS- VOTER ID/NON CAND." ~ "FOOD", disbursement_description == "MILEAGE REIMBURSEMENT-VOTER ID/NON CAND." ~ "MILEAGE REIMBURSEMENT", disbursement_description == "CATERING FOR FUNDRAISING EVENT-NON CAND." ~ "FOOD",disbursement_description == "MILEAGE REIMBURSEMENT- VOTER ID/NON CAND.
" ~ "REIMBURSEMENT", disbursement_description == "ONLINE FUNDRAISING FEE - NON CAND." ~ "ONLINE FUNDRAISING FEE- NON CAND.", disbursement_description == "MILEAGE REIMBURSEMENT" ~ "REIMBURSEMENT", disbursement_description == "FOOD FOR STAFF" ~ "FOOD", disbursement_description == "REFUND OF  CONTRIBUTION" ~ "REFUNDS", disbursement_description == "REFUND OF 9.9.2020 DONATION" ~ "REFUNDS", disbursement_description == "SALARY" ~ "SALARIES", disbursement_description == "FOOD FOR FUNDRAISING EVENT-NON CAND." ~ "FOOD", disbursement_description == "ONLINE FUNDRAISING FEE NON CAND." ~ "ONLINE ADVERTISING-NON CAND.", TRUE ~ disbursement_description)) %>% 
  arrange(desc(disbursement_amount))
```

#Who did most of the Republican money go to? What was the reason for the transaction? 

The Michigan Republican Party gave most of its money to itself, spending $287673.18. While there is not a description on the transaction that tells why most of that money was spent, we do know that $75,480.0.of it was given to the party for "levin transfer for split bills." The party spent a lot of money on creative and production fundraising at both Unisource Direct LLC ($149630.05) and ARC Print Solutions ($69533.33 for creative production and $35588.72 for postage for fundraising). 

I also find it interesting that the top choice for advertising was online. They party spent $149,195.29 on online advertising with the company IMGE.It's also interesting that the state's Republican Party spent $40,000 on Voter ID registration software from Campaign Sidekick. Does this have to do with the national Republican Party's lack of trust in the voting system?  

The Democratic Party spent a lot of money on GOTV (Get out the vote advertising) for advertising. The recipient of this transaction was Facebook. I believe that Facebook advertising was the top choice for Democrats. So, both parties invested heavily in online advertising. 

```{r}
TopRepReciever <- CleanMiRepSpend %>% 
  group_by(recipient_name,disbursement_description, or_disbursement_description) %>% 
  summarise(total= sum(disbursement_amount)
) %>%
  arrange(desc(total)) %>% 
  distinct()

TopRepNoDesc <- CleanMiRepSpend %>% 
  group_by(recipient_name) %>% 
   summarise(total= sum(disbursement_amount)
) %>%
  arrange(desc(total)) %>% 
  distinct()
```

#What transaction type was most of the money spent on for the Republican Party?--used OpenRefine

16% of the Republican Party's total spending was spent on creative and production fundraising ($228,959.37). 11.7% of the party's total spending was spent on online advertising ($167,285.40). 

```{r}
TransTypeRep <- CleanMiRepSpend %>% 
  group_by(disbursement_description) %>% 
  summarise(total= sum(disbursement_amount)
) %>%mutate(percent = (total/1427744)*100) %>% 
  arrange(desc(total)) %>% 
  distinct()
```

#Who spent more overall, Democrats or Republicans? 

Ultimately, the Republican Party spent less money than the Democratic Party. Michigan's Republican Party spent $1427744 compared to the Democratic Party that spent $5026673. That's a $3,598,929 difference. 

# Total Democratic Spending

```{r}
OverSpendDem <- CleanMiDemSpend %>% 
  group_by(committee_name) %>% 
  summarise(total_amount = sum(disbursement_amount))
OverSpendDem
```

#Total Republican Spending

```{r}
OverSpendRep <- CleanMiRepSpend %>% 
  group_by(committee_name) %>% 
  summarise(total_amount = sum(disbursement_amount))
OverSpendRep
```
MAPS AND CHARTS:

#Map of states' donations to the Michigan Democratic Party:

https://datawrapper.dwcdn.net/Rn0di/5/

```{r}
DemConMap <- StateDem %>% 
  mutate(Percent = (total/10382351)*100)

write.csv(DemConMap, "StatesData/DemConMap.csv")



```

#Map of states' donations to the Michigan Republican Party:

https://datawrapper.dwcdn.net/cAhdt/1/

```{r}
RepConMap <- StateRep %>% 
mutate(Percent = (total/2520106)*100)

write.csv(RepConMap, "StatesData/RepConMap.csv")

```

#Bar chart that shows which type of enity type donated the most money to the Democratic Party. 

https://datawrapper.dwcdn.net/XG8nZ/1/ 

```{r}
DemEnityChart <- DonatingMostDem
  
  write.csv(DemEnityChart, "StatesData/DemEnityChart.csv")

```


