---
title: "pre_lab_01.Rmd"
author: "Derek Willis"
date: "2022-07-05"
output: html_document
---
```{r}

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen=999)
```

## About this notebook

This notebook contains code and explanatory text that your should review and run as you read through chapters 1 - 5 of the course textbook, "Data Journalism with R and the Tidyverse". Answer questions and edit the document as directed.

Running this notebook will help you understand key data analysis methods and concepts that you will put into practice during this week's lab. When you come across a new command or verb in R, you should add it to your class reference notebook file along with what it's used for.

When you are finished running the code in this notebook, you will push changes to your course GitHub repo, and upload the link to ELMS as instructed.

You also should update your class reference notebook with code examples and what they mean. When you come across a new concept or R verb, put it in your reference notebook along with a sentence about what it does. For example:

```{r}
library(tidyverse)
```

The library function loads an R library such as the tidyverse, providing access to the verbs and functions that library contains. You have to load a library before you can use its functions.

## Instructor in class demonstration outline

* Load this notebook, and provide overview of R, RStudio (text editor, console, environment window, file explorer), Rmarkdown data notebooks, the reference notebook.
* Work through the code below, chapter by chapter.
  * Chapter 3: explain the console and the concept of variables.
  * Chapter 3: explain packages, the Tidyverse, how to install them.
  * Chapter 6: explain loading libraries, the concept of functions, loading data, the importance of having questions to guide data analysis, and aggregation and summarization.

## Chapter 3

### Task 1: The console
**Task** In the console (look below), type 2+2 and hit enter. What does it return? Type your answer below.
**Answer:**4

### Task 2: Variables
**Task** Put the following code -- which stores the numeral 2 as a variable called number -- in the console and hit enter.    

```{r}
number <- 2
```

Then type this code to multiply the variable called number by itself.  

```{r}
number * number
```

What does it return? Type your answer below:

**Answer:**4

### Task 3: Install packages
**Task** In the console, run the following code to install the Tidyverse:

```{r}
install.packages('tidyverse')
```

**Task** Add that

Then run the following code to install the rmarkdown library

```{r}
install.packages('rmarkdown')
```

## Chapter 6

### Task 1: Load libraries
**Task** Run the following code in the gray-colored codeblock below -- not in the console -- to load the tidyverse library. To run the code, click the little green play button (left facing arrow) at the top right of the codeblock. In Rmarkdown data notebooks, we write code inside of codeblocks, and explanatory text in the white area outside of it.   

```{r}
# This is a comment inside of a codeblock. Comments doesn't run when we run the codeblock.
# Adding this command is a good way to start your class reference notebook.
library(tidyverse)
```

### Task 2: Load data
**Task** Load the Maryland WinRed contributions dataset by running the following codeblock.

```{r}
maryland_winred_contributions <- read_rds("pre_lab_01.rds")
```

Once it loads, look at the environment window at right in RStudio for maryland_winred_contributions. It's next to a small blue circle with a white arrow in it.  To the right of that, look at how many "obs" the dataset has ("obs" is short for observation, which is another way of saying rows or records). Look at how many variables it has (another way of saying columns or fields).  Write the number of observations and variables below.

**Answer**
54247 Observations and 24 variables 
### Task 3: Glimpse data
**Task** Run the following codeblock to use glimpse() to get a sense of the data, some of the values in each field/variable/column, and the data types of each field/variable/column.

```{r}
glimpse(maryland_winred_contributions)
```

In the space below, write the data type for the "name" column and the "amount" column (it's in light gray, offset by < >).  Also offer your best guess as to what the "name" column and the "amount" column represent.
**Answer** I'm not sure what's meant by data type. My guess is that they represent the amount of donations received by each candidate by contributors. 

### Task 4: Examine data "head"
**Task** Run the following codeblock to use head() to examine the first six rows of data. Explore the data by clicking the black arrow on the right side of the output table, allowing you to examine additional columns.
```{r}
head(maryland_winred_contributions)
```
In the space below, write the name of the business from the first row of the "name" column.
**Answer** Homemaker

### Task 5: Explore data with interactive dataframe
**Task** look at the environment window at right in RStudio for maryland_winred_contributions It's next to a small blue circle with a white arrow in it. Click on the maryland_winred_contributions -- not the blue circle. It will open the dataframe as an interactive table.  Scroll down to see additional rows.  Scroll left to see additional columns. Click on a column name to sort by that column.

### Task 6: Group and count
**Task** Run the following codeblock:

```{r}
maryland_winred_contributions %>%
  group_by(date) %>%
  summarise(
    count_contribs = n()
  )
```

Describe what you think this code is doing below.

**Answer**I think it is sorting the contributions by date.I can't figure out what's going on in the second column. It looks like the count contributions are going from in the thousands in the previous chart to the hundreds.The count contributions with the earliest dates are first.

### Task 7
**Task** Run the following codeblock:

```{r}
maryland_winred_contributions %>%
  group_by(date) %>%
  summarise(
    count_contribs = n()
  ) %>%
  arrange(desc(count_contribs))
```
Compared to the previous codeblock, what has changed and how did that affect the result?
**Answer**The contribution amounts are now in descending order. So now the contribution amount from 2022-06-29 is at the top of the list because it is the biggest. 

### Task 8
**Task** Run the following codeblock.
```{r}
maryland_winred_contributions %>%
  group_by(memo_text, amount) %>%
  summarise(
    count_contribs = n()
  ) %>%
  arrange(desc(count_contribs))
```

Compared to the previous codeblock, what has changed and how did that affect the result?
**Answer** The memo text was moved to the first column followed by the amount and count contributions. The count contributions are in descending order.

### Task 9
**Task** Run the following codeblock:
```{r}
maryland_winred_contributions %>%
  group_by(memo_text) %>%
  summarise(
    count_contribs = n(),
    total_amount = sum(amount)
  ) %>%
  arrange(desc(total_amount))
```
Compared to the previous codeblock, what has changed and how did that affect the result?
**Answer**Now the total amounts are in descending order. That's how this entire chart is organized. It changes the order that the memo text are in.

### Task 10
**Task** Run the following codeblock to group the maryland_winred_contributions dataframe by the recipient for each contribution. Then do the following summarizations for each group:count the number of contributions, add up the value of all contributions, and calculate the mean and median contribution amounts. Then sort so the recipient with the most contributions is listed first.
```{r}
maryland_winred_contributions %>%
  group_by(memo_text) %>%
  summarise(
    count_contribs = n(),
    total_amount = sum(amount),
    mean_amount = mean(amount),
    median_amount = median(amount)
  ) %>%
  arrange(desc(count_contribs))
```
Compared to the previous codeblock, what has changed and how did that affect the result? Glancing at the names, are any of the top results from Maryland?
**Answer**None of the top results are from Maryland. The first two results in the memo text are the same. The Republican National Committee moves to the third spot while the second NRCC falls to the fourth slot.Parrott For Congress and Ambrose fell completely out of the top slots.McCarthy, Johnson, and Rubio all switch spots. 

### Task 11
**Task** Run the following codeblock to group the maryland_winred_contributions dataframe by the city for each contribution. Then do the following summarizations for each group:count the number of contributions, add up the value of all contributions, calculate the mean and median contributions amounts and calculate the smallest and largest contributions Then sort so the city with the largest total contribution amount is listed first.

```{r}
maryland_winred_contributions %>%
  group_by(city) %>%
  summarise(
    count_contribs = n(),
    total_amount = sum(amount),
    mean_amount = mean(amount),
    median_amount = median(amount),
    min_amount = min(amount),
    max_amount = max(amount)
  ) %>%
  arrange(desc(total_amount))
```
Compared to the previous codeblock, what has changed and how did that affect the result? How is the data sorted?
**Answer**
The cites in Maryland where these contributions were provided are now the left most column. The total contribution amounts are in descending order. 
